How and Why to run a Hack Week

### 2017-09-20 sections/what.tex

Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
Hackathons originated from the open-source software movement, and have historically focused on software development and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
With this in mind, hackathons around scientific topics, designed to foster collaboration \cite{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \cite{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.

@@

Our past hack week events follow most closely the communal hackathon model, as it applies to scientific communities of practice.
Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure 1).
For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.

>>>>

Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
Hackathons originated from the open-source software movement, and have historically focused on software development, particularly in the form of coding sprints, and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
With this in mind, hackathons around scientific topics, designed to foster collaboration \cite{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \cite{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.

@@

Our past hack week events follow most closely the communal hackathon model, as it applies to scientific communities of practice.
Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.

### 2017-09-20 sections/why.tex

\begin{itemize}
\item{\textit{Education and Training}: 
While some hack weeks are focused more on education than others (Figure 1), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices. Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields. This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}

>>>>

\begin{itemize}
\item{\textit{Education and Training}: 
While some hack weeks are focused more on education than others, there is often a skill-development component that entails extensive discussion on reproducible research and open science practices. Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields. This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
